Unsupervised training of maximum-entropy models for lexical selection in rule-based machine translation

### 2015-04-12 2015-eamt-melsa/2015-eamt-melsa.tex
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59727 72bbbca6-d526-0410-a7d9-f06f51895060

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{multirow}
\usepackage{url}
\usepackage{natbib}
\usepackage{color}
\usepackage[small,bf]{caption}
\usepackage[backgroundcolor=lightgray]{todonotes}
\newcommand{\mytodo}[1]{\todo[inline]{#1}}
\newcommand{\todomlf}[1]{\renewcommand{\baselinestretch}{0.55}\todo{\scriptsize\sf MLF: #1}\renewcommand{\baselinestretch}{1.0}}

@@

\section{Introduction}
\todomlf{R1 has a general comment: ``watch out for long sentences''}
Corpus-based MT has been the primary research direction in the field
of machine translation in recent years. However, rule-based MT (RBMT)
systems are still being developed, and there are many successful
commercial and non-commercial systems. One reason for the continued
development of RBMT systems is that in order to be successful,
corpus-based MT requires parallel corpora in the order of tens of
millions of words. Although for some language pairs these exist, they
only exist for a fraction of the world's languages.

An RBMT system typically consists of an analysis
component,\footnote{Such as a morphological or syntactic analyser.} a
transfer component and a generation component. As part of the transfer
component it is necessary to make choices regarding words in the
source language which may have more than one translation in the target
language.

\emph{Lexical selection} is the task of choosing, for a given
source-language word, the most adequate translation in the target
language among a known set of alternatives.  The task is related to
the task of word-sense disambiguation \citep{ide98a}. However, it is
different to word-sense disambiguation in that lexical selection is a
bilingual problem, not a monolingual problem: its aim is to find the
most adequate translation, not the most adequate sense. Thus, it is
not necessary to choose between a series of fine-grained senses if all
these senses result in the same final translation; however, it may
sometimes be necessary to choose a different translation for the same
sense, for example in a collocation.
% It could also be the case that a single sense could have more than
% one possible translation (for example synonyms).

@@

\subsection{Prior work}
%@@@ \comment{This section should be ironed out a bit.}  
%In their pioneering work, 
\cite{dagan1994word} used the term \emph{word sense disambiguation} to
refer to what is actually target word selection in machine
translation; they used a parser to identify syntactic relations
such as subject--object or subject--verb.  After generating all the
possible translations for a given input sentence using an ambiguous
bilingual dictionary, they extract the syntactic tuples from the
target language and count the frequency in a previously-trained
target-language model of tuples.  They use maximum-likelihood
estimation to calculate the probability that a given target language
tuple is the translation of a given source language tuple, with an
automatically determined confidence threshold.

Later, \cite{berger1996} illustrated the use of maximum-entropy
classifiers on the specific problem of lexical selection in IBM-style
word-based statistical machine translation. Other authors
\citep{melero07a} have used target-language models to
rank %@@@ carbonell06
the translations resulting from all possible combinations of lexical
selections. Nowadays, in state-of-the-art phrase-based
statistical % \todo{METIS too?}
machine translation \citep{koehn2009statistical}, lexical selection is
taken care of by a combination of the translation model and the
language model. The translation model provides probabilities of
translation between words or word sequences (often referred to as
\emph{phrases}) in the source and target language. The target-language
model provides probabilities of word sequences in the target language.
% As regards rule-based machine translation, \cite{sanchez07a}
% proposed to use information from the target language to train a
% statistical lexical-selection module but did not provide any
% experiments.  @@@ \todo[inline]{FMT: integrate \citep{sanchez07a}
% somewhere here}
\cite{marechek10} trained a maximum-entropy lexical selector for their
dependency-grammar-based transfer system TectoMT using a bilingual corpus.
More recently, \cite{tyers12a} presented a method of lexical selection for rule-based
machine translation based on rules which select or remove translations
in fixed-length contexts, along with a training method for learning
the rules from a word-aligned parallel corpus.

@@

\section{Method}
\todomlf{R3 says: ``The technique clearly resembles the "Deciphering Foreign Language," (S. Ravi and K. Knight) or “EM Decipherment for Large Vocabularies “
(M. Nuhn, and H. Ney. EM) approaches, although these techniques were not meant for learning the disambiguation but for learning the translation dictionaries.
The proposed method only executes one EM iteration in contrast to these works. It would be interesting to highlight differences and resemblances with these deciphering approaches.''. Check and cite? Also, single iteration WTF?  }
%\todo[inline]{FSM: In the experiments a development corpus is
%  mentioned. For what is it used? This should be explain in this
%  section.}

%\comment{Improve this} 
Lexical selection in this paper considers for
each word a simple source-language context made up of neighbouring
lemma+part-of-speech combinations. Contexts considered include up to
two words to the left and up to two words to the right of the word to
be translated. 
%\comment{FSM: Later, a development corpus is
%  mentioned. In this section it should be mentioned what it is used
% for.}

Let the probability of a word $t$ being the translation of a
word $s$ in a source-language context $c$ be $p_s(t|c)$. In principle, this value
could be estimated directly from the available corpora for every
combination of $(s, t, c)$.  This would however present two questions:
(1) how should the relevant contexts be chosen? and (2) what should be
done when $(s, t, c)$ is not found in the corpus?
A maximum-entropy model answers both of these questions. It allows the
contexts that we consider to be linguistically interesting to be
defined \emph{a priori} and then integrate these seamlessly into a
probabilistic model \citep{Manning99b}.  In answer to the second
question, a maximum-entropy model assumes nothing about what is not
present in the training data. That is, if there is no information in
the training data, then it assumes that all outcomes ---that is, all
possible translations--- are equally likely.  As previously mentioned,
the principle of maximum entropy has been applied to the problem of
lexical selection before; in particular, \cite{berger1996} cast the
problem of lexical selection in statistical MT as a classification
problem. They learn a separate maximum-entropy classifier for each SL
word form,
%\comment{[:we do too, integrate this later]} 
using SL context to
distinguish between possible translations. These classifiers are then
incorporated into the translation model of their word-based SMT
system.
%\comment{[how is our work different?]}  
In their approach, a
classifier consists of a set of binary feature functions and
corresponding weights for each feature. 
In both \cite{berger1996} and our method, features are defined in the
form $h_k^{s}(t, c)$, where $t$
is a translation, and $c$ is a source-language (SL) context.\todomlf{Defined SL which had not been defined} One difference is that \cite{berger1996} take $s$, $t$
and $c$ to be based on word forms, whereas in our method they are based
on lemma forms.
An example would be the following feature where the Spanish word \emph{pez} (`fish' as a living animal) is seen as the
translation of \emph{arrain} (`fish') in the context \emph{arrain
  handi} `big fish' would therefore be defined as:
%\comment{FSM: I have introduced the index $k$
%  because several features are defined for a SL word $s$.} 
\begin{equation}
h_\mathrm{+handi}^{\mathrm{arrain}}(t, c) = \left \{ \begin{matrix}
                 ~~1 & \mbox{if } 
\left\{ \begin{array}{c}
t~ =~ \textit{pez}~ \\ \mathrm{and}~ \\ \textit{handi}~ \mathrm{follows}~ \textit{arrain} \end{array} \right. \\
                 ~~0 & \mbox{otherwise} \end{matrix} \right.
\label{eq:feat-func}
\end{equation}
This feature considers a context of zero words to the left of the
problem word and one word (+ \emph{handi}) to the right of it.\todomlf{Removed \(k\) and used a more informative subscript}

@@

\begin{equation}
p_s(t|c) = \frac{1}{Z^s(c)} \exp{ \sum_{k=1}^{n_F} \lambda_k^s h_k^s(t, c)}
\label{eq:max-ent-sum}
\end{equation}
\todomlf{R3 complains: `` Probability notation is confusing. I would suggest the use of \(p(t | c,s)\) instead of \(p_s(t|c)\), that will simplify notation. For instance instead of \(Z^s (c)\) we would have \(Z(c,s)\).''. Ignore or take onboard?}

@@

The input to our method consists of a collection of samples,
$\mathcal{G} = (S, G)$, where $S = (s_1, s_2, \ldots s_{|S|})$ is a
sequence of SL words, and $G = \{g_1, g_2, \ldots g_{|G|}\}$ is a set
of possible \emph{lexical-selection paths}. A lexical-selection path
$g = (t_1, t_2, \ldots, t_{|S|})$ is a sequence of lexical-selection
choices of those SL words, where \(t_i\) is an element of \(T_s(s_i)\), the set of possible translations of \(s_i\).\todomlf{R1 asked for this, even if redundant, and says: ``can one source word be translated as multiple words? What about multi-word expressions that cannot be translated word-by-word or source words that do not have a TL equivalent'', etc.} % and \(T_s(\cdot)\)
%is the set of all
%translations for a source word.
This is produced in the first stages of RBMT, just after morphological
analysis, part-of-speech tagging, and bilingual dictionary lookup, and
before any structural transfer takes place (we will call this
\emph{pre-lexsel}). In our model, it is after these first stages that
lexical selection (\emph{lexsel}) occurs. After lexical selection,
structural transfer and generation take place; a function
$\tau(g_i,S)$ represents the result of these last stages, which we
will call \emph{post-lexsel}, and returns a finished translation of a
specific lexical-selection path $g_i$ of sentence
$S$. Figure~\ref{fig:lexselschema} shows this process
schematically. %%%\todo{Can we try to bring the figure closer?}
\todomlf{R3 says: ``it would have been nicer to see an example of one sentence going through the pipeline''. Can we add this in the text?}
%Taking our previous example for the parallel corpus, Table~\ref{table:ex-monocorp} 
%presents an example of five possible samples.

@@

specific lexical selection events \((s,t,c)\) cannot be counted, a
target language model \(P_\mathrm{TL}(\cdot)\) is used to compute a value for
the fractional count for disambiguation path \(g_i\), $p(g_i|S)$ after
suitable normalisation:\todomlf{R2 says ``Nevertheless, I've got problems to understand (4), in particular the paths in post-lexsel step. Some better explanation could be offered maybe by saying that (4) evaluates all possible solutions in the language model.''. Don't know what to do here. }

@@

when in \(g_i\) $t$ is the translation for \(s\) in context
\(c\). That is, as if event $(s,t,c)$ had been seen a fractional
number $p(g_i|S)$ of times. We prune $(s, t, c)$
occurring less than a certain number of
times in the corpus, using a development corpus to guide pruning (see
section~\ref{sec:eval-ref-results}).

@@

part-of-speech tagger in a rule-based machine translation system.

@@

This section describes the training and evaluation settings used in
the remainder of this paper. The primary motivation behind the
evaluation is that it should be automatic, meaningful, and be performed over a test set which
is large enough to be representative. It should evaluate both
performance on the specific subtask of lexical selection, and on the
whole translation task. Evaluating lexical-selection performance is an
\emph{intrinsic} module-based evaluation.  It measures how well the lexical
selection module disambiguates the lexical-transfer
output as compared to a gold-standard corpus. The lexical 
transfer output is the result of looking up the translations of the 
source-language \emph{lexical forms} --- lemmas and  % combinations of lemma and tags
tags --- in the bilingual dictionary.

@@

\item[Breton--French:] Development of the Breton--French pair has been
  described in \citep{tyers10b}. The bilingual % tyers09b
  dictionaries were not built with polysemy in mind from the outset,
  but some entries were added later to start work on lexical
  selection. The version used in this paper is SVN revision
  41375.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-br-fr}}
%  This is the only available machine translation system between Breton
%  and French. It has been developed part-time over a number of years.
\todomlf{R1 says ``whole section is rather length especially the revision numbers and URLs of the specific translations systems need not be mentioned in my opinion. The space could rather be used to elaborate in the description of the method in Sec 2.''}

\item[Macedonian--English:] The Macedonian--English pair
  in Apertium was created specifically for the purposes of running
  lexical-selection experiments. 
% The resources reused from other pairs
%  were the English morphological dictionary from the Icelandic and
%  English pair \citep{brandt11}, the Macedonian morphological
%  dictionary and constraint grammar from the Macedonian and Bulgarian
%  pair \citep{rangelov11}, and the SETimes parallel corpus
  The lexical resources for the pair were tuned to the SETimes parallel corpus
  \citep{tyers10}.  
% The work was carried out over a period of eight
%  days, and consisted of creating the bilingual dictionary and
%  transfer rules. The bilingual dictionary was created by tagging both
%  sides of the parallel corpus, word-aligning them with {\sc giza++}
%  \citep{och03a} and extracting the probabilistic lexicon. Entries
%  from this lexicon were checked manually according to frequency and
%  included in the bilingual dictionary of the machine translation
%  system. 
  The most probable entry from automatic word alignment of this corpus
  using {\sc giza++} \citep{och03a} was checked to ensure that it was
  an adequate translation, and if so marked as the default.\footnote{A
    bilingual dictionary in Apertium \citep{forcada2011apertium} may
    contain several target-language translation for a given SL
    word. Where there is more than one possible translation, the
    dictionary writer chooses the most general or most frequent
    translation among the set of possible translations and marks it as
    the \emph{linguistic default} translation.} As a result of
  attempting to include all possible translations, the average number
  of translations per word is much
  higher % (see table~\ref{table:lang-pair-stats})
  than in other pairs.
% The transfer rules were written
%  by hand to treat translation problems in the corpus. 
The version of
  the software used in this paper is SVN revision
  41476.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-mk-en}}
\item[Basque--Spanish:] The development of the Basque to Spanish pair
  is described in \cite{ginesti09}. 
% The bilingual dictionary was taken
%  from the free/open-source Matxin system \citep{mayor2011}. 
  Alternative translations were included in the bilingual
  dictionary. The version used is SVN revision 44846.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-eu-es}}
% , but non-default translations were marked with a
%  direction restriction\comment{FSM: This needs to be carefully
%    explained to avoid confusing the reader.}.\footnote{The bilingual
%    dictionaries in Apertium are bidirectional. They contain entries
%    which can be used for translating both from language $A$ to
%    language $B$ and vice versa. A direction restriction is used to
%    limit the entry to being used for translation in one translation
%    direction.} For lexical-selection experiments these direction
%  restrictions were removed.

\item[English--Spanish:] The English--Spanish pair was developed from
  a combination of the English--Catalan and Spanish--Catalan pairs, and
%  over a period of around four months. The pair
%has been used in other
%  lexical selection experiments \citep{sanchez07a}\comment{FSM: Sure?
%    It was on en-ca and in that paper no experimental results are
%    shown. This paper would be better mentioned in the introduction
%    (it mentions the use of TL info for lexical selection.},
  contains a number of entries in the bilingual dictionary with more
  than one translation. %The most-frequent translation, as judged by
%  the language pair developer, is marked as the \emph{default}
%  translation and non-default translations were added according to
%  frequency. \todo{Default is explained for some LPs only, why?}  
  The
  version used in this paper is SVN revision
  41387.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-en-es}}\todomlf{4-month mention removed in response to R1}
\end{description}

@@

For creating the test corpora, providing a source language corpus 
for training, and a target language corpus for scoring, we used four parallel corpora: 
\begin{itemize}
\addtolength{\itemsep}{-0.10in}
\item \textbf{Ofis ar Brezhoneg} (\texttt{OAB}): This parallel corpus
  of Breton and French has been collected specifically for
  lexical-selection experiments from translations produced by
  \emph{Ofis ar Brezhoneg} `The Office of the Breton
  language'. %\footnote{In 2010 the administrative status was changed
%    and it was renamed \emph{Ofis Publik ar Brezhoneg} `Public office
%    of the Breton language'. The corpus was published before this date
%    so we use the original name.}  
%  It contains some parallel data
%  previously described in \cite{tyers09b}, and also some new data. 
The corpus has recently been made available online through {\sc opus}.\footnote{\url{http://opus.lingfil.uu.se}}
\item \textbf{South-East European Times} (\texttt{SETimes}): Described
  in \cite{tyers10}, this corpus is a multilingual corpus of the
  Balkan languages (and English) in the news domain. The Macedonian and English
  part will be used.%\todo{Not in OPUS?}
\item \textbf{Open Data Euskadi} (\texttt{OpenData}): This is a
  Basque--Spanish parallel corpus made from the translation
  memories of the \emph{Herri Arduralaritzaren Euskal Erakundea}
  `Basque Institute of Public
  Administration'.\footnote{\url{http://tinyurl.com/eu-es-tm}}
\item \textbf{European Parliament Proceedings} (\texttt{EuroParl}):
  Described by \cite{koehn05a}, this is a multilingual corpus of the
  European Union official languages. We are using the English--Spanish
  data from version 7.\footnote{\url{http://www.statmt.org/europarl/}}
\end{itemize}

>>>>

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{multirow}
\usepackage{url}
\usepackage{natbib}
\usepackage{color}
\usepackage[small,bf]{caption}
\usepackage[backgroundcolor=lightgray]{todonotes}
\newcommand{\mytodo}[1]{\todo[inline]{#1}}
\newcommand{\todomlf}[1]{\renewcommand{\baselinestretch}{0.55}\todo{\scriptsize\sf MLF: #1}\renewcommand{\baselinestretch}{1.0}}
\newcommand{\todofsm}[1]{\renewcommand{\baselinestretch}{0.55}\todo{\scriptsize\sf FSM: #1}\renewcommand{\baselinestretch}{1.0}}

@@

\section{Introduction}
\todomlf{R1 has a general comment: ``watch out for long sentences''}
Corpus-based machine translation (MT) has been the primary research
direction in the field of MT in recent years. However, rule-based MT
(RBMT) systems are still being developed, and there are many
successful commercial and non-commercial systems. One reason for the
continued development of RBMT systems is that in order to be
successful, corpus-based MT requires parallel corpora in the order of
tens of millions of words. Although for some language pairs these
exist, they only exist for a fraction of the world's languages.

An RBMT system typically consists of an analysis
component,\footnote{Such as a morphological or syntactic analyser.} a
transfer component and a generation component. As part of the transfer
component it is necessary to make choices regarding words in the
source language (SL) which may have more than one translation in the
target language (TL).

\emph{Lexical selection} is the task of choosing, for a given SL word,
the most adequate translation in the TL among a known set of
alternatives.  The task is related to the task of word-sense
disambiguation \citep{ide98a}. However, it is different to word-sense
disambiguation in that lexical selection is a bilingual problem, not a
monolingual problem: its aim is to find the most adequate translation,
not the most adequate sense. Thus, it is not necessary to choose
between a series of fine-grained senses if all these senses result in
the same final translation; however, it may sometimes be necessary to
choose a different translation for the same sense, for example in a
collocation.
% It could also be the case that a single sense could have more than
% one possible translation (for example synonyms).

@@

\subsection{Prior work}
%@@@ \comment{This section should be ironed out a bit.}  
%In their pioneering work, 
\cite{dagan1994word} used the term \emph{word sense disambiguation} to
refer to what is actually target word selection\todofsm{Should we call
  it lexical selection instead of target word selection?} in MT; they
used a parser to identify syntactic relations such as subject--object
or subject--verb.  After generating all the possible translations for
a given input sentence using an ambiguous bilingual dictionary, they
extract the syntactic tuples from the TL and count the frequency in a
previously-trained TL model of tuples.  They use maximum-likelihood
estimation to calculate the probability that a given TL tuple is the
translation of a given SL tuple, with an automatically determined
confidence threshold.

Later, \cite{berger1996} illustrated the use of maximum-entropy
classifiers on the specific problem of lexical selection in IBM-style
word-based statistical MT. Other authors \citep{melero07a} have used
TL models to rank %@@@ carbonell06
the translations resulting from all possible combinations of lexical
selections. Nowadays, in state-of-the-art phrase-based
statistical % \todo{METIS too?}
MT \citep{koehn2009statistical}, lexical selection is taken care of by
a combination of the translation model and the language model. The
translation model provides probabilities of translation between words
or word sequences (often referred to as \emph{phrases}) in the source
and target language. The TL model provides probabilities of word
sequences in the TL.
% As regards rule-based machine translation, \cite{sanchez07a}
% proposed to use information from the target language to train a
% statistical lexical-selection module but did not provide any
% experiments.  @@@ \todo[inline]{FMT: integrate \citep{sanchez07a}
% somewhere here}
\cite{marechek10} trained a maximum-entropy lexical selector for their
dependency-grammar-based transfer system TectoMT using a bilingual
corpus.  More recently, \cite{tyers12a} presented a method of lexical
selection for RBMT based on rules which select or remove translations
in fixed-length contexts, along with a training method for learning
the rules from a word-aligned parallel corpus.

@@

\section{Method}
\todomlf{R3 says: ``The technique clearly resembles the "Deciphering
  Foreign Language," (S. Ravi and K. Knight) or “EM Decipherment for
  Large Vocabularies “ (M. Nuhn, and H. Ney. EM) approaches, although
  these techniques were not meant for learning the disambiguation but
  for learning the translation dictionaries.  The proposed method only
  executes one EM iteration in contrast to these works. It would be
  interesting to highlight differences and resemblances with these
  deciphering approaches.''. Check and cite? Also, single iteration
  WTF?}
%\todo[inline]{FSM: In the experiments a development corpus is
%  mentioned. For what is it used? This should be explain in this
%  section.}

%\comment{Improve this} 
Lexical selection in this paper considers for each word a simple SL
context made up of neighbouring lemma+part-of-speech
combinations. Contexts considered include up to two words to the left
and up to two words to the right of the word to be translated.
%\comment{FSM: Later, a development corpus is
%  mentioned. In this section it should be mentioned what it is used
% for.}

Let the probability of a word $t$ being the translation of a word $s$
in a SL context $c$ be $p_s(t|c)$. In principle, this value could be
estimated directly from the available corpora for every combination of
$(s, t, c)$.  This would however present two questions: (1) how should
the relevant contexts be chosen? and (2) what should be done when $(s,
t, c)$ is not found in the corpus?  A maximum-entropy model answers
both of these questions. It allows the contexts that we consider to be
linguistically interesting to be defined \emph{a priori} and then
integrate these seamlessly into a probabilistic model
\citep{Manning99b}.  In answer to the second question, a
maximum-entropy model maximise the entropy subject to match the
expected counts of the designed features with those found in the
training data. \todofsm{Solved R3: ``In answer to the second question, a
  maximum-entropy model assumes nothing about what is not present in
  the training data. That is, if there is no information in the
  training data, then it assumes that all outcomes —that is, all
  possible translations— are equally likely'' can be applied to a
  uniform distribution for a discrete variable. I think the authors
  are trying to restate the main property of the maximum entropy
  model, i.e.  maximum entropy model maximize the entropy (which is
  the uniformness of the distribution) subject to match the expected
  counts of the designed features with those found in the training
  data.  }
%assumes nothing about what is not present in the
%training data. 
That is, if there is no information in the training data, then it
assumes that all outcomes ---that is, all possible translations--- are
equally likely.  As previously mentioned, the principle of maximum
entropy has been applied to the problem of lexical selection before;
in particular, \cite{berger1996} cast the problem of lexical selection
in statistical MT as a classification problem. They learn a separate
maximum-entropy classifier for each SL word form,
%\comment{[:we do too, integrate this later]} 
using SL context to distinguish between possible translations. These
classifiers are then incorporated into the translation model of their
word-based statistical MT system.
%\comment{[how is our work different?]}  
In their approach, a classifier consists of a set of binary feature
functions and corresponding weights for each feature.  In both
\cite{berger1996} and our method, features are defined in the form
$h_k^{s}(t, c)$, where $t$ is a translation, and $c$ is a SL
context. One difference is that \cite{berger1996} take $s$, $t$ and
$c$ to be based on word forms, whereas in our method they are based on
lemma forms.  An example would be the following feature where the
Spanish word \emph{pez} (`fish' as a living animal) is seen as the
translation of \emph{arrain} (`fish') in the context \emph{arrain
  handi} `big fish' and would therefore be defined as:
%\comment{FSM: I have introduced the index $k$
%  because several features are defined for a SL word $s$.} 
\begin{equation}
h_\mathrm{+handi}^{\mathrm{arrain}}(t, c) = \left \{ \begin{matrix}
                 ~~1 & \mbox{if } 
\left\{ \begin{array}{c}
t~ =~ \textit{pez}~ \\ \mathrm{and}~ \\ \textit{handi}~ \mathrm{follows}~ \textit{arrain} \end{array} \right. \\
                 ~~0 & \mbox{otherwise} \end{matrix} \right.
\label{eq:feat-func}
\end{equation}
This feature considers a context of zero words to the left of the
problem word and one word (+ \emph{handi}) to the right of it.\todomlf{Removed \(k\) and used a more informative subscript}

@@

\begin{equation}
p_s(t|c) = \frac{1}{Z^s(c)} \exp{ \sum_{k=1}^{n_F} \lambda_k^s h_k^s(t, c)}
\label{eq:max-ent-sum}
\end{equation}
\todomlf{R3 complains: `` Probability notation is confusing. I would
  suggest the use of \(p(t | c,s)\) instead of \(p_s(t|c)\), that will
  simplify notation. For instance instead of \(Z^s (c)\) we would have
  \(Z(c,s)\).''. Ignore or take onboard? FSM: Mention that we follow
  the notation of Berger?}

@@

The input to our method consists of a collection of samples,
$\mathcal{G} = (S, G)$, where $S = (s_1, s_2, \ldots s_{|S|})$ is a
sequence of SL words, and $G = \{g_1, g_2, \ldots g_{|G|}\}$ is a set
of possible \emph{lexical-selection paths}. A lexical-selection path
$g = (t_1, t_2, \ldots, t_{|S|})$ is a sequence of lexical-selection
choices of those SL words, where \(t_i\) is an element of \(T_s(s_i)\), the set of possible translations of \(s_i\).\todomlf{R1 asked for this, even if redundant, and says: ``can one source word be translated as multiple words? What about multi-word expressions that cannot be translated word-by-word or source words that do not have a TL equivalent'', etc.} % and \(T_s(\cdot)\)
%is the set of all
%translations for a source word.
This is produced in the first stages of RBMT, just after morphological
analysis, part-of-speech tagging, and bilingual dictionary lookup, and
before any structural transfer takes place (we will call this
\emph{pre-lexsel}). In our model, it is after these first stages that
lexical selection (\emph{lexsel}) occurs. After lexical selection,
structural transfer and generation take place; a function
$\tau(g_i,S)$ represents the result of these last stages, which we
will call \emph{post-lexsel}, and returns a finished translation of a
specific lexical-selection path $g_i$ of sentence
$S$. Figure~\ref{fig:lexselschema} shows this process
schematically. %%%\todo{Can we try to bring the figure closer?}
\todomlf{R3 says: ``it would have been nicer to see an example of one
  sentence going through the pipeline''. Can we add this in the text?
  FSM: We do not have much space}
%Taking our previous example for the parallel corpus, Table~\ref{table:ex-monocorp} 
%presents an example of five possible samples.

@@

specific lexical selection events \((s,t,c)\) cannot be counted, a TL
model \(P_\mathrm{TL}(\cdot)\) is used to compute a value for the
fractional count for disambiguation path \(g_i\), $p(g_i|S)$ after
suitable normalisation:\todomlf{R2 says ``Nevertheless, I've got
  problems to understand (4), in particular the paths in post-lexsel
  step. Some better explanation could be offered maybe by saying that
  (4) evaluates all possible solutions in the language model.''. Don't
  know what to do here. }

@@

when in \(g_i\) the translation for \(s\) in context \(c\) is
$t$. That is, as if event $(s,t,c)$ had been seen a fractional number
$p(g_i|S)$ of times. We prune $(s, t, c)$ occurring less than a
certain number of times in the corpus, using a development corpus to
guide pruning (see section~\ref{sec:eval-ref-results}).

@@

part-of-speech tagger in a RBMT system.

@@

This section describes the training and evaluation settings used in
the remainder of this paper. The primary motivation behind the
evaluation is that it should be automatic, meaningful, and be performed over a test set which
is large enough to be representative. It should evaluate both
performance on the specific subtask of lexical selection, and on the
whole translation task. Evaluating lexical-selection performance is an
\emph{intrinsic} module-based evaluation.  It measures how well the lexical
selection module disambiguates the lexical-transfer
output as compared to a gold-standard corpus. The lexical 
transfer output is the result of looking up the translations of the 
SL \emph{lexical forms} --- lemmas and  % combinations of lemma and tags
tags --- in the bilingual dictionary.

@@

\item[Breton--French:] Development of the Breton--French pair has been
  described in \citep{tyers10b}. The bilingual % tyers09b
  dictionaries were not built with polysemy in mind from the outset,
  but some entries were added later to start work on lexical
  selection.\footnote{The version used in this paper is SVN revision
  41375; \url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-br-fr}}
%  This is the only available machine translation system between Breton
%  and French. It has been developed part-time over a number of years.
  \todomlf{R1 says ``whole section is rather length especially the
    revision numbers and URLs of the specific translations systems
    need not be mentioned in my opinion. The space could rather be
    used to elaborate in the description of the method in Sec
    2.''. FSM: I have moved them to the footnote}

\item[Macedonian--English:] The Macedonian--English pair
  in Apertium was created specifically for the purposes of running
  lexical-selection experiments. 
% The resources reused from other pairs
%  were the English morphological dictionary from the Icelandic and
%  English pair \citep{brandt11}, the Macedonian morphological
%  dictionary and constraint grammar from the Macedonian and Bulgarian
%  pair \citep{rangelov11}, and the SETimes parallel corpus
  The lexical resources for the pair were tuned to the SETimes
  parallel corpus \citep{tyers10}.
% The work was carried out over a period of eight
%  days, and consisted of creating the bilingual dictionary and
%  transfer rules. The bilingual dictionary was created by tagging both
%  sides of the parallel corpus, word-aligning them with {\sc giza++}
%  \citep{och03a} and extracting the probabilistic lexicon. Entries
%  from this lexicon were checked manually according to frequency and
%  included in the bilingual dictionary of the machine translation
%  system. 
  The most probable entry from automatic word alignment of this corpus
  using {\sc giza++} \citep{och03a} was checked to ensure that it was
  an adequate translation, and if so marked as the default.\footnote{A
    bilingual dictionary in Apertium \citep{forcada2011apertium} may
    contain several TL translation for a given SL word. Where there is
    more than one possible translation, the dictionary writer chooses
    the most general or most frequent translation among the set of
    possible translations and marks it as the \emph{linguistic
      default} translation.} As a result of attempting to include all
  possible translations, the average number of translations per word
  is much higher % (see table~\ref{table:lang-pair-stats})
  than in other pairs.\footnote{The version of the software used in this paper is SVN revision
  41476; \url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-mk-en}}
\item[Basque--Spanish:] The development of the Basque to Spanish pair
  is described in \cite{ginesti09}. 
% The bilingual dictionary was taken
%  from the free/open-source Matxin system \citep{mayor2011}. 
  Alternative translations were included in the bilingual
  dictionary.\footnote{The version used is SVN revision 44846; \url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-eu-es}}
% , but non-default translations were marked with a
%  direction restriction\comment{FSM: This needs to be carefully
%    explained to avoid confusing the reader.}.\footnote{The bilingual
%    dictionaries in Apertium are bidirectional. They contain entries
%    which can be used for translating both from language $A$ to
%    language $B$ and vice versa. A direction restriction is used to
%    limit the entry to being used for translation in one translation
%    direction.} For lexical-selection experiments these direction
%  restrictions were removed.

\item[English--Spanish:] The English--Spanish pair was developed from
  a combination of the English--Catalan and Spanish--Catalan pairs, and
%  over a period of around four months. The pair
%has been used in other
%  lexical selection experiments \citep{sanchez07a}\comment{FSM: Sure?
%    It was on en-ca and in that paper no experimental results are
%    shown. This paper would be better mentioned in the introduction
%    (it mentions the use of TL info for lexical selection.},
  contains a number of entries in the bilingual dictionary with more
  than one translation.\footnote{The
  version used in this paper is SVN revision
  41387; \url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-en-es}}

%The most-frequent translation, as judged by
%  the language pair developer, is marked as the \emph{default}
%  translation and non-default translations were added according to
%  frequency. \todo{Default is explained for some LPs only, why?}  
\todomlf{4-month mention removed in response to R1}
\end{description}

@@

For creating the test corpora, providing a SL corpus for training, and
a TL corpus for scoring, we used four parallel corpora:
\begin{itemize}
\addtolength{\itemsep}{-0.10in}
\item \textbf{Ofis ar Brezhoneg} (\texttt{OAB}): This parallel corpus
  of Breton and French has been collected specifically for
  lexical-selection experiments from translations produced by
  \emph{Ofis ar Brezhoneg} `The Office of the Breton
  language'. %\footnote{In 2010 the administrative status was changed
%    and it was renamed \emph{Ofis Publik ar Brezhoneg} `Public office
%    of the Breton language'. The corpus was published before this date
%    so we use the original name.}  
%  It contains some parallel data
%  previously described in \cite{tyers09b}, and also some new data. 
The corpus has recently been made available online through {\sc opus}.\footnote{\url{http://opus.lingfil.uu.se}}
\item \textbf{South-East European Times} (\texttt{SETimes}): Described
  in \cite{tyers10}, this corpus is a multilingual corpus of the
  Balkan languages (and English) in the news domain. The Macedonian and English
  part will be used.%\todo{Not in OPUS?}
\item \textbf{Open Data Euskadi} (\texttt{OpenData}): This is a
  Basque--Spanish parallel corpus made from the translation
  memories of the \emph{Herri Arduralaritzaren Euskal Erakundea}
  `Basque Institute of Public
  Administration'.\footnote{\url{http://tinyurl.com/eu-es-tm}}
\item \textbf{European Parliament Proceedings} (\texttt{EuroParl}):
  Described by \cite{koehn05a}, this is a multilingual corpus of the
  European Union official languages. We are using the English--Spanish
  data from version 7.\footnote{\url{http://www.statmt.org/europarl/}}
\end{itemize}

There are a number of approaches to creating evaluation corpora for
lexical selection in the literature. \cite{vickrey05} use a parallel
corpus to make annotated test and training sets for experiments in
lexical selection applied to a simplified translation problem in
statistical MT. They use word alignments from {\sc giza++}
\citep{och03a} to annotate SL words with their translations from the
reference translation in the parallel corpus. One disadvantage of this
method is that only one translation is annotated per SL word, meaning
that accuracies may be lower because of missing translations --- this
happens when the system chooses a translation which is adequate, but
is not found in the reference translation. A second disadvantage is
that the word alignments may not be 100\% reliable, which decreases
the accuracy of the annotated corpus. An alternative method is
described by \cite{zinovjeva2000}, who manually tags ambiguous words
in English sentences with their % a selection of
translation in Swedish. %The size of the annotated corpus is not
%mentioned. 
%\cite{specia2005a} use a parallel corpus, but eschew using
%word alignment in favour of using a bilingual dictionary and heuristic
%rules based on part-of-speech and relative position. The corpus was
%reviewed manually before being used in further experiments
%\citep{specia2005}.
